Road traffic sound level estimation from realistic urban sound mixtures by Non-negative Matrix Factorization

### 2018-03-01 paper/gloaguenAppliedNMF.tex
minor improvements

$d_{\beta}$ is mainly chosen as a $\beta$-divergence \cite{fevotte_algorithms_2011}, a sub-classes belonging to the Bregman divergences \cite{hennequin_beta-divergence_2011} which include 3 specific divergence and distance calculations: the Euclidean distance (eq. \ref{eq:def_distEUC}), the Kullback-Leibler divergence (eq. \ref{eq:def_divKL}) and the Itakura-Sa\"{i}to divergence (eq. \ref{eq:def_divIS}):

@@

\subsection{Semi-supervised NMF}
The main issue with the supervised approach is the limit imposed by $\mathbf{W}$: to be successful, all the acoustical sources must be considered in it which is not possible in an urban environment. As the main source that interests us can be known \textit{a priori}, semi-supervised NMF (SEM-NMF) \cite{lee_semi-supervised_2010} is considered to better take into account the interfering sound sources. This approach proposes to decompose $\mathbf{W}_{F \times (K+J)}$ with two distinctive matrices: $\mathbf{W} = \left[ \mathbf{W_s}~\mathbf{W_r} \right]$ where $\mathbf{W_s}_{F \times K}$ is a fixed part of $\mathbf{W}$ composed of fixed audio spectra and $ \mathbf{W_r}_{F \times J}$, a mobile part which is updated, see eq. \ref{eq:W_r_SS}. Thus it is possible to include elements not present in $\mathbf{W_s}$. The dimension of $\mathbf{W_r}$ is set up as $J << K$ in order to best considered the sound source present in $\mathbf{W_s}$. $\mathbf{H}$ is then also decomposed in two matrices, $\mathbf{H}_{(K+J) \times N} = \genfrac[]{0pt}{0}{\mathbf{H_s}}{\mathbf{H_r}}$. The eq. \ref{eq:nmf} becomes

\begin{equation}
\mathbf{V} = \mathbf{W_s H_s}+\mathbf{W_r H_r}.
\end{equation}

@@

This approach as the advantage, with the add of the mobile part $\mathbf{W_r}$, to bring more flexibility and then to be more adaptive to the different urban sound environments. Applications of Sem-NMF can be found for musical \cite{weninger2012supervised, kitamura_music_2014} and speech content \cite{joder2012real, mysore2011non}.

@@

The urban sound scenes are taken from 76 recordings from 2 to 5 min, achieved in the \nth{13} district of Paris (France) at 19 different locations \footnote{Recordings were made as part of the Grafic project funded by Ademe}, which cover four various sound environments (Figure \ref{fig:map_grafic}). A complete description of the experimental protocol can be found in \cite{aumond_modelling_2017}. Two of the 76 recordings are rejected for the analysis because the audio files were corrupted, resulting in 74 valid audio files assumed as representative of the variety of sound environments. The recordings are listened and categorized within four different sound environments, as proposed in \cite{can_describing_2015}: park (P, 8 audio files), quiet street (Q, 35 audio files), noisy street (N, 23 audio files) and very noisy street (vN, 8 audio files). Then, each audio file is annotated, noting the start and end time of each sound event along with its sound class (described here below). The aim of the annotation phase is next to transcribe the recordings, in order to obtain simulated sound scenes with the same distribution of sound events as the recordings and therefore as close as possible to the realistic scenes.

@@

First the spectrogram of each audio file is computed (window $w = 2^{12}$ with 50 $\%$ overlap). The spectrogram is then cut in multiple temporal frames of $w_t  = \lbrace$0.5, 1$\rbrace$ second duration. In each cut spectrogram, the root mean square on each frequency bin is calculated to obtained a spectrum of $F \times 1$ dimension. This method allows the description of the audio sample with finer spectra and then having the different characteristic pitches of the traffic spectra. An illustrative example on a 3 seconds sample is displayed in Figure \ref{fig:example_dictionary}. From the 53 audio files, we obtain respectively for $w_t\in$ $\lbrace$0.5, 1$\rbrace$ second, 2218 and 1109 elements. A $K$-mean clustering algorithm is applied to reduce these dimensions to $K \in \lbrace$25, 50, 100, 200$\rbrace$ in order to avoid redundant information and decrease the computation time. The obtained $K$ clusters compose then the $K$ elements of $\mathbf{W}$. Furthermore, the case where each audio generates one spectrum from its spectrogram is added ($w_t = all$). Here, by the added approach, the dictionary elements are based on the spectral envelops of the audio samples. For this case, having 53 audio samples, the number of elements in $\mathbf{W}$ with the $K$-mean clustering algorithm is reduced to $K \in \lbrace$25, 50$\rbrace$.

@@

\begin{tabularx}{17.5cm}{L{3cm}@{}C{4cm}@{}@{}@{}C{4cm}@{}@{}C{4cm}@{}C{2cm}@{}}	
	\rowcolor[HTML]{C0C0C0} 
  \textbf{method} & Sup NMF & Sem NMF & TI NMF & 3\\
\end{tabularx}

@@

The LP filter with $f_c = 20$ kHz cut-off frequency is equivalent to considerate the sound level of the entire scene without specific distinction between the sound sources. The error is then important with a high standard deviation ($mMAE =$ 3.76 ($\pm$ 4.35) dB). The lowest error for a LP filter is obtained with $f_c$ = 500 Hz ($mMAE = $2.14 ($\pm$ 1.83) dB). It allows a balance between the discarded and remaining energy through the sound environments.

@@

Furthermore, on the dictionary creation, except SEM-NMF where all the best methods according to $\beta$ use the same dictionary, no specific dictionary form through all the method, is used. SUP and SEM-NMF privilege a high number of element ($K$ = 200) with a fine description of the audio samples ($w_t \in \lbrace 0.5$,$1 \rbrace$ second). For TI-NMF, the composition of the initial dictionary is more diverse both on the $K$ value and on the finesse  of the description ($w_t \in \lbrace all$,$0.5 \rbrace$ second).

@@

Except SEM-NMF, all the methods show the same error evolution the same error evolution: a decrease of the error with the increase of the traffic predominance. SEM-NMF show an almost constant error for all 4 sound environments. The LP filter error is mainly important for environments where the traffic is less present. As this approach considers the remaining energy as the traffic component, no distinction can be made between the different sound sources not related to the sound sources of interest, here traffic. In the opposite, for noisy and very noisy environments, the performances of the LP filter are good ($MAE$ $<$ 1 dB). The errors are then due to a high deletion of the traffic energy by the filter while it becomes the main sound source.

@@

Besides having the lowest $mMAE$ error, TI-NMF presents the most performing results. The park environment is the only case where an other NMF approach out-performed significantly firm TI-NMF ($MAE$ = 3.05 dB). In this sound environment, the traffic dictionary is then composed, in average, of more than half the 200 elements in $\mathbf{W'}$ (136 elements). For the rest of the sound environments, firm TI-NMF has the lowest error. For very noisy environment the error is even very low ($MAE=$ 0.30 dB). In this case, in average, $\mathbf{W'}$ is composed of 198 traffic elements. This method out-performed SUP-NMF since, as $\mathbf{W_0}$ is updated, $\mathbf{W'}$ is then most suited to the sound scene than a fixed dictionary. The advantage to have an unique dictionary suited to each sound scene makes TI-NMF very performing when traffic is predominant while, by the thresholding, it limits the dictionary derivation when the traffic is more quiet.

@@

A non-negative matrix factorization frame has been applied as a source separation tool to estimate the traffic sound level  from a corpus of urban sound scenes, which consists of scenes built artificially in which the level and position of each source is controled. In addition, the realism of the scenes has been verified thanks to a perceptual test. As a result, the proposed protocol mimics the outputs given by a deployed sensor network, but knowing the actual sound level of each source composing the sound mixture.

>>>>

$d_{\beta}(x\vert y)$ is mainly chosen as a $\beta$-divergence \cite{fevotte_algorithms_2011}, a sub-classes belonging to the Bregman divergences \cite{hennequin_beta-divergence_2011} which include 3 specific divergence and distance calculations: the Euclidean distance (eq. \ref{eq:def_distEUC}), the Kullback-Leibler divergence (eq. \ref{eq:def_divKL}) and the Itakura-Sa\"{i}to divergence (eq. \ref{eq:def_divIS}):

@@

\subsection{Semi-supervised NMF}
The main issue with the supervised approach is the limit imposed by $\mathbf{W}$: to be successful, all the acoustical sources must be considered in it which is not possible in an urban environment. As the main source that interests us can be known \textit{a priori}, semi-supervised NMF (SEM-NMF) \cite{lee_semi-supervised_2010} is considered to better take into account the interfering sound sources. This approach proposes to decompose $\mathbf{W}_{F \times (K+J)}$ with two distinctive matrices: $\mathbf{W} = \left[ \mathbf{W_s}~\mathbf{W_r} \right]$ where $\mathbf{W_s}_{F \times K}$ is a fixed part of $\mathbf{W}$ composed of audio spectra and $ \mathbf{W_r}_{F \times J}$, a mobile part which is updated, see eq. \ref{eq:W_r_SS}. Thus it is possible to include elements not present in $\mathbf{W_s}$. The dimension of $\mathbf{W_r}$ is set up as $J << K$ in order to best considered the sound source present in $\mathbf{W_s}$. $\mathbf{H}$ is then also decomposed in two matrices, $\mathbf{H}_{(K+J) \times N} = \genfrac[]{0pt}{0}{\mathbf{H_s}}{\mathbf{H_r}}$. The eq. \ref{eq:nmf} becomes

\begin{equation}
\mathbf{V} \approx \mathbf{WH} = \mathbf{W_s H_s}+\mathbf{W_r H_r}.
\end{equation}

@@

This approach as the advantage, with the add of the mobile part $\mathbf{W_r}$, to bring more flexibility and then to be more adaptive to the different urban sound environments. Applications of SEM-NMF can be found for musical \cite{weninger2012supervised, kitamura_music_2014} and speech content \cite{joder2012real, mysore2011non}.

@@

The urban sound scenes are taken from 76 recordings from 2 to 5 min, achieved in the \nth{13} district of Paris (France) at 19 different locations \footnote{Recordings were made as part of the Grafic project funded by Ademe}, which cover four various sound environments (Figure \ref{fig:map_grafic}), see Figure \ref{fig:map_grafic}. A complete description of the experimental protocol can be found in \cite{aumond_modelling_2017}. Two of the 76 recordings are rejected for the analysis because the audio files were corrupted, resulting in 74 valid audio files assumed as representative of the variety of sound environments. The recordings are listened and categorized within four different sound environments, as proposed in \cite{can_describing_2015}: park (P, 8 audio files), quiet street (Q, 35 audio files), noisy street (N, 23 audio files) and very noisy street (vN, 8 audio files). Then, each audio file is annotated, noting the start and end time of each sound event along with its sound class. The aim of the annotation phase is next to transcribe the recordings, in order to obtain simulated sound scenes with the same distribution of sound events as the recordings and therefore as close as possible to the realistic scenes.

@@

First the spectrogram of each audio file is computed (window $w = 2^{12}$ with 50 $\%$ overlap). The spectrogram is then cut in multiple temporal frames of $w_t  = \lbrace$0.5, 1$\rbrace$ second duration. In each cut spectrogram, the root mean square on each frequency bin is calculated to obtained a spectrum of $F \times 1$ dimension. This method allows the description of the audio sample with finer spectra and then having the different characteristic pitches of the traffic spectra. An illustrative example on a 3 seconds sample is displayed in Figure \ref{fig:example_dictionary}. From the 53 audio files, we obtain respectively for $w_t\in$ $\lbrace$0.5, 1$\rbrace$ second, 2218 and 1109 elements. A $K$-mean clustering algorithm is applied to reduce these dimensions to $K \in \lbrace$25, 50, 100, 200$\rbrace$ in order to avoid redundant information and decrease the computation time. The obtained $K$ clusters compose then the elements of $\mathbf{W}$. Furthermore, the case where each audio generates one spectrum from its spectrogram is added ($w_t = all$). Here, by the added approach, the dictionary elements are based on the spectral envelops of the audio samples. For this case, having 53 audio samples, the number of elements in $\mathbf{W}$ with the $K$-mean clustering algorithm is reduced to $K \in \lbrace$25, 50$\rbrace$.

@@

\begin{tabularx}{17.5cm}{L{3cm}@{}C{4cm}@{}@{}@{}C{4cm}@{}@{}C{4cm}@{}C{2cm}@{}}	
	\rowcolor[HTML]{C0C0C0} 
  \textbf{method} & SUP NMF & SEM NMF & TI NMF & 3\\
\end{tabularx}

@@

The LP filter with $f_c = 20$ kHz cut-off frequency is equivalent to considerate the sound level of the entire scene without specific distinction between the sound sources. The error is then important with a high standard deviation ($mMAE =$ 3.76 ($\pm$ 4.35) dB). The lowest error for a LP filter is obtained with $f_c$ = 500 Hz ($mMAE =$ 2.14 ($\pm$ 1.83) dB). It allows a balance between the discarded and remaining energy through the sound environments.

@@

Furthermore, on the dictionary creation, except SEM-NMF where all the best methods according to $\beta$ use the same dictionary, no specific dictionary form through all the method, is used. SUP and SEM-NMF privilege a high number of element ($K$ = 200) with a fine description of the audio samples ($w_t \in \lbrace 0.5$,$1 \rbrace$ second). For TI-NMF, the composition of the initial dictionary is more diverse both on the $K$ value and on the finesse  of the description ($w_t \in \lbrace all$,$0.5$, $1 \rbrace$ second).

@@

Except SEM-NMF, all the methods show the same error evolution the same error evolution: a decrease of the error with the increase of the traffic predominance. SEM-NMF show an almost constant error for all 4 sound environments. The LP filter error is mainly important for environments where the traffic is less present. As this approach considers the remaining energy as the traffic component, no distinction is made between the different sound sources not related to the traffic component. In the opposite, for noisy and very noisy environments, the performances of the LP filter are good ($MAE$ $<$ 1 dB). The errors are then due to a high deletion of the traffic energy by the filter while it becomes the main sound source.

@@

Besides having the lowest $mMAE$ error, TI-NMF presents the most performing results. The park environment is the only case where an other NMF approach out-performed significantly firm TI-NMF ($MAE$ = 3.05 dB). In this sound environment, the traffic dictionary is then composed, in average, of more than half the 200 elements in $\mathbf{W'}$ (136 elements). For the rest of the sound environments, firm TI-NMF has the lowest error. For very noisy environment the error is even very low ($MAE=$ 0.30 dB). In this case, in average, $\mathbf{W'}$ is composed of 198 traffic elements. This method out-performed SUP-NMF since, as $\mathbf{W_0}$ is updated, the final dictionary $\mathbf{W'}$ is then most suited to the sound scene than a fixed dictionary. The advantage to have an unique dictionary suited to each sound scene makes TI-NMF very performing when traffic is predominant while, by the thresholding, it limits the dictionary derivation when the traffic is less present.

@@

A non-negative matrix factorization frame has been applied as a source separation tool to estimate the traffic sound level  from a corpus of urban sound scenes artifically built. In addition, the realism of the scenes has been verified thanks to a perceptual test. The proposed protocol generate audio sample similar than the outputs of a deployed sensor networks with tha advantages of the simulation process (sound level and position of each source controlled and known).
